A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility

### 2019-05-09 paper.Rmd
Typo fixes

```{r fig-effort-underrstanding, fig.cap = "There is a big difference in the effort to prepare data, and how easy it is to udnerstand - look at the difference between most datasets, and something like a Randomized Control Trial (RCT)."}
library(ggplot2)

@@

A more common use case is to include data as part of a regular package where analytical methods are being implemented and shared. This serves the dual purpose of exposing methods and data together, making it extremely easy for a researcher to simply install and load the library at the top of their script. CRAN's distributed network (along with the global content distribution network maintained by Rstudio) ensure that the package is quickly accessible to anyone in the R ecosystem. A second critical advantage in this approach is that one could also benefit from R's package documentation syntax to generate useful metadata for fast reference. This approach can also easily be adapted to other languages. Python for example, is far more flexible with respect to including arbitrary files as part of a package.

@@

One major disadvantage of packaging data inside R is that it makes the data availability very language centric. Non R users are unlikely to download and export data out of a package. This is why we recommend, as a rule, that researchers also archive data in a long-term data repository. These include domain specific repositories (see SECTION) or more general purpose ones such as Zenodo or figshare and include the persistent identifier in all locations where the data is referenced such as the manuscript, notebook and data package.

@@

1. Database access [@data-access] for services like MySQL, Hive, and ODBC, similar to R’s DBI organisation [@R-DBI]
2. Accessing data online via APIs, such as accessing the weather source NOAA [@NOAA-jl]
3. Data is provided directly. For example, accessing Econometric datasets [@econ-data-jl], face datasets [@face-data-jl], Market data [@market-data-jl], and even R datasets [@r-data-jl]. It is telling that the datasets provided in R are useful, as there is even a Julia package, RDatasets that provides over 700 datasets from R packages. There are also packages for Accessing datasets used in vega [@vega-data-jl], for accessing common machine learning datasets [@ml-data-jl]. There are also packages for generating fake data [@fake-data-jl], for which there are at least two R packages that do this, charlatan [@charlatan-data-r], and [@wakefield-data-r].

@@

In **Domain Specific Venue**, data can be published in journal data papers, or venues specific to the type of data. For example, in astronomy the data is hosted at the SDSS (SDSS), Genetic data can be hosted at GenBank (@GenBank), and so on.  Some other places to put data include the open science framework: https://osf.io/, dataverse, https://dataverse.org/, and data.world, https://data.world/. Other datasets might be produced

@@

This contains fields such as path, and a schema with subfields name and type, for each variable. It also provides information for, licensing and features such as line breaks, and delimiters. It is built on JSON (JavaScript Object Notation), a lightweight, human-friendly, machine readable data-interchange format. Table schema is baked into formats such as [csvy](http://csvy.org/), an extended csv format, which has additional front matter in a YAML format using Table Schema.

@@

The data used in the data analysis should be provided in a folder called `data`.
Ideally, the data should be in "Tidy Data" format [@Wickham2014], where tidy data contains variables in columns, and observations in rows. Contrasting `raw data`, `tidy data`/`analysis data` should be in an easily readable plain-text format, such as CSV, tab separated, or semicolon separated. Binary or proprietary formats are discouraged in favour of interoperability.

@@

This is an enourmous, ongoing research effort. Brief context of the data is provided at the link, along with a link to the census plot data from 2012 at [this paper](https://repository.si.edu/handle/10088/20925). This provides an abstract and description of the data, along with co-authors, and a recommended acknowledgement.

@@

The SDSS provides different degrees of understanding, for different understanding. It is a huge project involving many people, and so would have provided enourmous effort to prepare, but has a high degree of understanding. It would be in the top right hand corner of the DARECO.

>>>>

```{r fig-effort-underrstanding, fig.cap = "There is a big difference in the effort to prepare data, and how easy it is to understand - look at the difference between most datasets, and something like a Randomized Control Trial (RCT)."}
library(ggplot2)

@@

A more common use case is to include data as part of a regular package where analytical methods are being implemented and shared. This serves the dual purpose of exposing methods and data together, making it extremely easy for a researcher to simply install and load the library at the top of their script. CRAN's distributed network (along with the global content distribution network maintained by RStudio) ensure that the package is quickly accessible to anyone in the R ecosystem. A second critical advantage in this approach is that one could also benefit from R's package documentation syntax to generate useful metadata for fast reference. This approach can also easily be adapted to other languages. Python for example, is far more flexible with respect to including arbitrary files as part of a package.

@@

One major disadvantage of packaging data inside R is that it makes the data availability very language centric. Non R users are unlikely to download and export data out of a package. This is why we recommend, as a rule, that researchers also archive data in a long-term data repository. These include domain specific repositories (see SECTION) or more general purpose ones such as Zenodo or Figshare and include the persistent identifier in all locations where the data is referenced such as the manuscript, notebook and data package.

@@

1. Database access [@data-access] for services like MySQL, Hive, and ODBC, similar to R’s DBI organization [@R-DBI]
2. Accessing data online via APIs, such as accessing the weather source NOAA [@NOAA-jl]
3. Data is provided directly. For example, accessing Econometric datasets [@econ-data-jl], face datasets [@face-data-jl], Market data [@market-data-jl], and even R datasets [@r-data-jl]. It is telling that the datasets provided in R are useful, as there is even a Julia package, RDatasets that provides over 700 datasets from R packages. There are also packages for Accessing datasets used in vega [@vega-data-jl], for accessing common machine learning datasets [@ml-data-jl]. There are also packages for generating fake data [@fake-data-jl], for which there are at least two R packages that do this, charlatan [@charlatan-data-r], and [@wakefield-data-r].

@@

In **Domain Specific Venue**, data can be published in journal data papers, or venues specific to the type of data. For example, in astronomy the data is hosted at the SDSS (SDSS), Genetic data can be hosted at GenBank (@GenBank), and so on.  Some other places to put data include the open science framework: https://osf.io/, Dataverse, https://dataverse.org/, and data.world, https://data.world/. Other datasets might be produced

@@

This contains fields such as path, and a schema with subfields name and type, for each variable. It also provides information for, licensing and features such as line breaks, and delimiters. It is built on JSON (JavaScript Object Notation), a lightweight, human-friendly, machine readable data-interchange format. Table schema is baked into formats such as [csvy](http://csvy.org/), an extended `csv` format, which has additional front matter in a YAML format using Table Schema.

@@

The data used in the data analysis should be provided in a folder called `data`.
Ideally, the data should be in "Tidy Data" format [@Wickham2014], where tidy data contains variables in columns, and observations in rows. Contrasting `raw data`, `tidy data`/`analysis data` should be in an easily readable plain-text format, such as CSV, tab separated, or semicolon separated. Binary or proprietary formats are discouraged in favor of interoperability.

@@

This is an enormous, ongoing research effort. Brief context of the data is provided at the link, along with a link to the census plot data from 2012 at [this paper](https://repository.si.edu/handle/10088/20925). This provides an abstract and description of the data, along with co-authors, and a recommended acknowledgement.

@@

The SDSS provides different degrees of understanding, for different understanding. It is a huge project involving many people, and so would have provided enormous effort to prepare, but has a high degree of understanding. It would be in the top right hand corner of the DARECO.
